
BOOK ONE

The Doctrine of Being

I.40
paramanu-parama-mahattvanto 'sya vasikara

GENERAL DIVISION OF BEING

Being is determined, first, as against another in general;
secondly, it is internally self-determining;
thirdly, as this preliminary division is cast off,
it is the abstract indeterminateness and immediacy
in which it must be the beginning.

According to the first determination,
being partitions itself off from essence,
for further on in its development it proves to be
in its totality only one sphere of the concept,
and to this sphere as moment it opposes another sphere.
According to the second, it is the sphere
within which fall the determinations
and the entire movement of its reflection.

In this, being will posit itself in three determinations:

I. as determinateness; as such, quality;

II. as sublated determinateness; magnitude, quantity;

III. as qualitatively determined quantity; measure.

This division, as was generally remarked of
such divisions in the Introduction,
is here a preliminary statement;
its determinations must first arise
from the movement of being itself,
and receive their definitions and justification by virtue of it.
As regards the divergence of this division from
the usual listing of the categories,
namely quantity, quality, relation and modality,
(for Kant, incidentally, these are supposed to be
only classifications of his categories,
but are in fact themselves categories,
only more abstract ones;
about this, there is nothing to remark here,
since the entire listing will diverge from
the usual ordering and meaning of
the categories at every point.)

This only can perhaps be remarked,
that the determination of quantity
is ordinarily listed ahead of quality and as a rule
this is done for no given reason.
It has already been shown that
the beginning is made with being as such,
and hence with qualitative being.
It is clear from a comparison of quality with quantity
that the former is by nature first.
For quantity is quality which has already become negative;
magnitude is the determinateness which,
no longer one with being but already distinguished from it,
is the sublated quality that has become indifferent.
It includes the alterability of being
without altering the fact itself,
namely being, of which it is the determination;
qualitative determinateness is on the contrary one with its being,
it neither transcends it nor stays within it
but is its immediate restrictedness.
Hence quality, as the determinateness which is immediate,
is the first and it is with it that the beginning is to be made.

Measure is a relation, not relation in general
but specifically of quality and quantity to each other;
the categories dealt with by Kant under relation
will come up elsewhere in their proper place.
Measure, if one so wishes, can be considered also a modality;
but since with Kant modality is no longer
supposed to make up a determination of content,
but only concerns the reference of the content
to thought, to the subjective,
the result is a totally heterogeneous reference
that does not belong here.

The third determination of being falls within
the section Quality inasmuch as being, as abstract immediacy,
reduces itself to one single determinateness
as against its other determinacies inside its sphere.

SECTION I

Determinateness (Quality)

Being is the indeterminate immediate;
it is free of determinateness with respect to essence,
just as it is still free of any determinateness
that it can receive within itself.
This reflectionless being is being
as it immediately is only within.

Since it is immediate, it is being without quality;
but the character of indeterminateness attaches to it in itself
only in opposition to what is determinate or qualitative.
Determinate being thus comes to stand over and against being in general;
with that, however, the very indeterminateness of being
constitutes its quality.
It will therefore be shown that the first being is
in itself determinate, and therefore secondly,
that it passes over into existence, is existence;
that this latter, however, as finite being, sublates itself
and passes over into the infinite reference of being to itself;
it passes over, thirdly, into being-for-itself.

CHAPTER 1

Being

I.41
ksina-vrtter abhijatasyeva maner
grahitr-grahana-grahyesu tat-stha-tad-anjanata samapatti

A. BEING

Being, pure being, without further determination.
In its indeterminate immediacy it is equal only to itself
and also not unequal with respect to another;
it has no difference within it, nor any outwardly.
If any determination or content were posited in it as distinct,
or if it were posited by this determination or content
as distinct from an other,
it would thereby fail to hold fast to its purity.
It is pure indeterminateness and emptiness.
There is nothing to be intuited in it,
if one can speak here of intuiting;
or, it is only this pure empty intuiting itself.
Just as little is anything to be thought in it,
or, it is equally only this empty thinking.
Being, the indeterminate immediate is in fact nothing,
and neither more nor less than nothing.

B. NOTHING

Nothing, pure nothingness;
it is simple equality with itself,
complete emptiness,
complete absence of determination and content;
lack of all distinction within.
In so far as mention can be made here of
intuiting and thinking,
it makes a difference whether something or nothing is
being intuited or thought.
To intuit or to think nothing has therefore a meaning;
the two are distinguished and so nothing is (concretely exists)
in our intuiting or thinking;
or rather it is the empty intuiting and thinking itself,
like pure being.
Nothing is therefore the same determination
or rather absence of determination,
and thus altogether the same as what pure being is.

C. BECOMING

1. Unity of being and nothing

Pure being and pure nothing are therefore the same.
The truth is neither being nor nothing,
but rather that being has passed over into nothing
and nothing into being;
“has passed over,” not passes over.

But the truth is just as much that
they are not without distinction;
it is rather that they are not the same,
that they are absolutely distinct
yet equally unseparated and inseparable,
and that each immediately vanishes in its opposite.

Their truth is therefore this movement of
the immediate vanishing of the one into the other:
becoming, a movement in which the two are distinguished,
but by a distinction which has just as immediately dissolved itself.

2. The moments of becoming

Becoming is the unseparatedness of being and nothing,
not the unity that abstracts from being and nothing;
as the unity of being and nothing
it is rather this determinate unity,
or one in which being and nothing equally are.
However, inasmuch as being and nothing are
each unseparated from its other, each is not.
In this unity, therefore, they are,
but as vanishing, only as sublated.
They sink from their initially represented self-subsistence
into moments which are still distinguished
but at the same time sublated.

Grasped as thus distinguished,
each is in their distinguishedness
a unity with the other.
Becoming thus contains being and nothing as two such unities,
each of which is itself unity of being and nothing;
the one is being as immediate and as reference to nothing;
the other is nothing as immediate and as reference to being;
in these unities the determinations are of unequal value.

Becoming is in this way doubly determined.
In one determination, nothing is the immediate,
that is, the determination begins with nothing
and this refers to being;
that is to say, it passes over into it.
In the other determination, being is the immediate,
that is, the determination begins with being
and this passes over into nothing:
coming-to-be and ceasing-to-be.

Both are the same, becoming,
and even as directions that are so different
they interpenetrate and paralyze each other.
The one is ceasing-to-be;
being passes over into nothing,
but nothing is just as much the opposite of itself,
the passing-over into being, coming-to-be.
This coming-to-be is the other direction;
nothing goes over into being,
but being equally sublates itself
and is rather the passing-over into nothing;
it is ceasing-to-be.
They do not sublate themselves reciprocally
[the one sublating the other externally]
but each rather sublates itself in itself
and is within it the opposite of itself.

3. Sublation of becoming

The equilibrium in which coming-to-be and ceasing-to-be are poised
is in the first place becoming itself.
But this becoming equally collects itself in quiescent unity.
Being and nothing are in it only as vanishing;
becoming itself, however, is only by virtue of their being distinguished.
Their vanishing is therefore the vanishing of becoming,
or the vanishing of the vanishing itself.
Becoming is a ceaseless unrest that collapses into a quiescent result.

This can also be expressed thus:
becoming is the vanishing of being into nothing,
and of nothing into being,
and the vanishing of being and nothing in general;
but at the same time it rests on their being distinct.
It therefore contradicts itself in itself,
because what it unites within itself is self-opposed;
but such a union destroys itself.

This result is a vanishedness, but it is not nothing;
as such, it would be only a relapse into one of
the already sublated determinations
and not the result of nothing and of being.
It is the unity of being and nothing
that has become quiescent simplicity.
But this quiescent simplicity is being,
yet no longer for itself but as determination of the whole.

Becoming, as transition into
the unity of being and nothing,
a unity which is as existent
or has the shape of the one-sided
immediate unity of these moments,
is existence.

CHAPTER 2

Existence

I.42
tatra sabda-artha-jnana-vikalpa sankirna savitarka samapatti

Existence is determinate being;
its determinateness is existent determinateness, quality.
Through its quality, something is opposed to an other;
it is alterable and finite,
negatively determined not only towards an other,
but absolutely within it.
This negation in it,
in contrast at first
with the finite something,
is the infinite;
the abstract opposition
in which these determinations appear
resolves itself into oppositionless infinity,
into being-for-itself.

The treatment of existence is therefore in three divisions:

A. existence as such
B. something and other, finitude
C. qualitative infinity.

Transition

Ideality can be called the quality of the infinite;
but it is essentially the process of becoming,
and hence a transition, like the transition
of becoming into existence.
We must now explicate this transition.
This immanent turning back, as the sublating of finitude,
of finitude as such and equally of the negative finitude
that only stands opposite to it, is only negative finitude,
is self-reference, being.
Since there is negation in this being, the latter is existence;
but, further, since the negation is essentially
negation of the negation, self-referring negation,
it is the existence that carries the name of being-for-itself.

CHAPTER 3

Being-for-itself

I.43
smrti-parisuddha svarupa-sunya-iva-artha-matra-nirbhasa nirvitarka

In being-for-itself,
qualitative being is brought to completion;
it is infinite being;
the being of the beginning is void of determination;
existence is sublated but only immediately sublated being;
it thus contains, to begin with,
only the first negation, itself immediate;
being is of course retained as well,
and the two are united in existence in simple unity;
for this reason, however, each is in itself still unlike the other,
and their unity is still not posited.
Existence is therefore the sphere of differentiation,
of dualism, the domain of finitude.
Determinateness is determinateness as such;
being which is relatively, not absolutely, determined.
In being-for-itself, the distinction
between being and determinateness,
or negation, is posited and equalized.
Quality, otherness, limit, as well as reality, in-itselfness,
ought, and so forth, are the incomplete configurations of negation in being
which are still based on the differentiation of the two.
But since in finitude negation has passed over into infinity,
in the posited negation of negation,
negation is simple self-reference
and in it, therefore, the equalization with being:
absolutely determinate being.

First, being-for-itself is immediately
an existent-for-itself, the one.

Second, the one passes over into a multiplicity of ones,
repulsion or the otherness of the one
which sublates itself into its ideality, attraction.

Third, we have the alternating
determination of repulsion and attraction
in which the two sink into a state of equilibrium;
and quality, driven to a head in being-for-itself,
passes over into quantity.

SECTION II

Magnitude (Quantity)

I.44
etaya-iva savicara nirvicara ca suksma-visaya vyakhyata

I.45
suksma-visayatvam calinga-paryavasanam

I.46
ta eva sabija samadhi

The difference between quantity and quality has been indicated.
Quality is the first, immediate determinateness.
Quantity is the determinateness that
has become indifferent to being;
a limit which is just as much no limit;
being-for-itself which is absolutely
identical with being-for-another:
the repulsion of the many ones
which is immediate non-repulsion,
their continuity.

Because that which exists for itself is now so posited
that it does not exclude its other
but rather affirmatively continues in it,
it is then otherness, inasmuch as
existence surfaces again on this continuity
and its determinateness is at the same time
no longer simple self-reference,
no longer the immediate determinateness
of the existent something,
but is posited as repelling itself from itself,
as referring to itself in the determinateness
rather of an other existence
(a being which exists for itself);
and since they are at the same time indifferent limits,
reflected into themselves and unconnected,
determinateness is as such outside itself,
an absolute externality and a something just as external;
such a limit, the indifference of the limit as limit
and the indifference of the something to the limit,
constitutes the quantitative determinateness of the something.

In the first place, we have to distinguish pure quantity
from quantity as determinate, from quantum.
First, pure quantity is real being-for-itself
turned back into itself, with as yet no determinateness in it:
a compact, infinite unity which continues itself into itself.

Second, this quantity proceeds to determinateness,
and this is posited in it as a determinateness
that at the same time is none, is only external.
Quantity becomes quantum.
Quantum is indifferent determinateness,
that is, one that transcends itself, negates itself;
as this otherness of otherness, it lapses into infinite progress.
Infinite quantum, however, is sublated indifferent determinateness:
it is the restoration of quality.

Third, quantum in qualitative form is quantitative ratio.
Quantum transcends itself only in general;
in the ratio, however, it transcends itself into its otherness,
in such a way that this otherness in which it has its determination
is at the same time posited, is another quantum.
With this we have quantum as turned back into itself
and referring to itself as into its otherness.

At the foundation of this relation there still
lies the externality of quantum;
it is indifferent quanta that
relate themselves to each other,
that is, they have the reference
that mutually connects them
in this being-outside-itself.
The ratio is, therefore, only
a formal unity of quality and quantity,
and dialectic is its transition into
their absolute unity, in measure.

SECTION III

Measure

I.47
nirvicara-vaisaradye 'dhyatma-prasada

I.48
rtambhara tatra prajna

I.49
sruta-anumana-prajnabhyam anyavisaya visesa-arthatvat

I.50
taj-ja samskaro 'nya-samskara-pratibandhi

I.51
tasyapi nirodhe sarva-nirodhan nirbija samadhi

Abstractly expressed, quality and quantity are in measure united.
Being as such is the immediate equality of determinateness with itself.
This immediacy of determinateness has sublated itself.
Quantity is being that has returned to itself in such a way
that it is a simple self-equality indifferent to determinateness.
But this indifference is only the externality of having
the determinateness not in itself but in an other.
As third, we now have the externality that refers itself to itself;
as self-reference, it is at the same time sublated externality and
carries the difference from itself in it;
a difference which, as externality, is the quantitative moment,
and, as taken back into itself, the qualitative.

Since among the categories of transcendental idealism
modality comes after quantity and quality,
with relation inserted in between,
this is an appropriate place to say something about it.
In transcendental idealism, this category has the meaning that
it is the connection of the subject matter to thought.
As understood by that idealism, thought is as such
essentially external to the thing-in-itself.
Hence, inasmuch as the other categories have
the transcendental determination of
belonging only to consciousness,
but as its objective moment,
so modality, which is the category of
the connection to the subject,
possesses the determination of
reflection in itself in a relative sense,
that is to say, the objectivity
which is granted to the other categories
is lacking in those of modality;
these, according to Kant's words,
do not add in the least to the concept
as a determination of the object
but only express its relation to the faculty of cognition
(Cr. of Pure R., 2nd edn, pp. 99, 266).
The categories which Kant groups under modality
(possibility, actuality, and necessity)
will come up later in their proper place.
Kant did not apply the form of triplicity
(an infinitely important form even though with him
it occurred only as a formal spark of light)
to the genera of his categories (to quantity, quality, etc.),
but only to their species to which he also gave the name of genera.
He was therefore unable to hit upon the third to quality and quantity.

With Spinoza, the mode is likewise the third after substance and attribute;
Spinoza defines it as the affections of substance,
or as that which is in another through which it is also comprehended.
In this way of conceiving it, this third is externality as such;
as has already been mentioned, with Spinoza generally,
the rigidity of substance lacks the turning back into itself.

The remark just made extends to any of the systems of pantheism
which thought has in one way or other produced.
Being, the one, substance, the infinite, essence, is the first;
opposite this abstraction is the second which
can be mustered in an equally abstract form,
as is habitually done as the next step in any purely formalistic thinking,
namely all determinateness generally taken as the mere finite,
the mere accidental, the transitory, the extraneous and unessential, etc.
But the bond connecting this second with the first is too invasive
for the second not to be not equally grasped with the first;
thus with Spinoza the attribute is the whole of substance,
though as comprehended by the understanding,
which is itself a restriction of substance or mode;
and so the mode, the insubstantial as such which can
be grasped only through an other,
constitutes the opposite extreme of substance, the third.
Also Indian pantheism, taken abstractly, has attained
in its monstrous fantasies this refinement
which runs like a moderating thread across its excesses
as its one point of interest namely
that Brahma, the one of abstract thought,
progresses through the shape of Vishnu,
particularly in the form of Krishna, to the third, Shiva.
The determination of this third is that of mode, alteration,
coming-to-be and passing-away;
it is the field of externality in general.
This Indian trinity has tempted a comparison with the Christian,
and one must indeed acknowledge a common element in them.
But it is essential to be aware of the difference that separates them.
It is not just that this difference is infinite
but that the true infinity makes the difference.
The determination of the Indian third principle is that
it is the dispersal of the substantial unity into its opposite,
not its turning back to itself,
a spiritual void rather, not spirit.
In the true trinity, there is not only unity but unification;
the syllogism is brought to a unity which is full of content and actual,
a unity which in its totally concrete determination is spirit.
The principle of the mode and of alteration
does not, of course, exclude unity altogether.
In Spinozism, for instance, precisely the mode is as such untrue
while substance alone is what truly is;
everything is supposed to be reduced to substance,
and this is then a sinking of all content into
an only formal unity void of content.
As for Shiva, it too is again the great whole,
not distinct from Brahma, from Brahma itself, that is,
the distinction and the determinateness just disappear
without being preserved, without being sublated,
and the unity does not become concrete unity,
nor is the diremption reconciled.
The supreme goal of the human being,
relegated as he is to the sphere
of coming-to-be and passing-away,
of modality in general,
is to sink into unconsciousness,
into unity with Brahma, annihilation;
the Buddhist Nirvana, Nibbana, etc., is the same.

Now although the mode is as such abstract externality,
indifference to qualitative as well as quantitative determinations,
and nothing in the essence should depend on the external,
the unessential, it is nevertheless conceded that
in the many all depends on the how;
but this is to concede that the mode itself
essentially belongs to the substance of a thing,
a very indefinite connection but
one which at least implies that the externality of
the mode is not all that abstract an externality after all.
Here the mode has the definite meaning of being measure.
The Spinozistic mode, just like the Indian principle of alteration,
is the measureless.
The Greeks were aware that everything has a measure.
Parmenides himself introduces necessity after abstract being,
as the ancient limit which is imposed on all.
And this awareness, although still vague, is
the beginning of a much higher concept
than is contained in substance
and in the distinction of the mode from it.

Measure in its more developed, more reflected form is necessity.
Fate, Nemesis, ultimately comes down to a determination of measure.
Whatever renders itself beyond the pale, becomes too great, too high,
is brought down to the other extreme of being reduced to nothing,
so that the mean of measure, the medium, is restored.
That the Absolute, God, is the measure of all things, is not
a stronger statement of pantheism than the definition,
“the Absolute, God, is being,” but is infinitely truer.
Measure is indeed an external way of things, a more or less,
but one which is as at the same time reflected into itself,
not merely an indifferent and external determinateness
but one which exists in itself;
thus it is the concrete truth of being.
For this reason the nations have revered in it
the presence of something inviolable and sacred.

Already present in measure is the idea of essence,
namely of being self-identical in the immediacy of being determined,
so that this immediacy is reduced through the self-identity to
something mediated, just as the self-identity is equally mediated
only through this externality, but the mediation is one with itself:
this is reflection, the moments of which indeed are,
but in this being are absolutely nothing but
moments of their negative unity.
In measure, the qualitative element is quantitative;
the determinateness or the difference is indifferent
and therefore a difference which is none;
it is sublated and this quantitativeness,
as an immanent turning back in which it is qualitative,
constitutes the being-in-and-for-itself which is essence.
But measure is essence only implicitly in itself or in its concept;
this concept of measure is not yet posited.
Measure is as such still the existent unity of the
qualitative and the quantitative element;
its moments are an existence, a quality and some quanta of this quality
which, in themselves, are indeed only indivisible,
but do not yet have the meaning of this reflected determination.
In the development of measure, these moments are differentiated
but at the same time referred to each other,
so that the identity which they are in themselves becomes
their connection explicitly, that is, is posited.
The meaning of this development is the realization of measure
in which the latter posits itself in relation to itself
and consequently as moment as well;
through this mediation, measure is determined as sublated;
its immediacy as well as that of its moments disappears;
these moments are reflected and thus measure,
having disclosed what it is according to its concept,
has passed over into essence.

Measure is at first the immediate unity of
the qualitative and the quantitative element, so that it is,

first, a quantum that has qualitative meaning and is as measure.
As so implicitly determined in itself, its further determination is that
the difference of its moments,
of its qualitatively and quantitatively determined being,
is disclosed in it.
These moments further determine themselves into
wholes of measure which as such are self-subsistent,
and, inasmuch as they refer to each other essentially,
measure becomes,

second, a ratio of specific quanta, each an independent measure.
But their self-subsistence also rests essentially
on a quantitative relation and a difference of magnitude,
and so the self-subsistence becomes a transition
of one measure into another.
The result is that measure collapses into the measureless.
But this beyond of measure is the negativity of measure
only in itself; thus,

third, the indifference of the determinations of
measure is thereby posited, and measure
(real measure because of the negativity contained within it)
is posited as an inverse ratio of measures
which, as self-subsistent qualities,
essentially rest on only their quantity
and their negative reference to each other,
and consequently prove to be only moments of
their truly self-subsistent unity.
This unity is the reflection-into-itself of each
and the positing of them; it is essence.

The development of measure,
which we have attempted in what follows,
is among the most difficult of subject matters.
Starting with immediate, external measure,
it would have to proceed, on the one hand,
to the further abstract determination of
the quantitative aspect of natural things
(of a mathematics of nature);
on the other side, it would have to indicate the link
between this determination of measure
and the qualitative aspect of those things,
at least in general, for the detailed demonstration of the
link between the qualitative and the quantitative aspects
as they originate in the concept of a concrete object
belongs to the particular science of the concrete
(examples of which, concerning the law of falling bodies
and the free movement of the heavens,
will be found in the Encyclopedia of the Philosophical Sciences).
We may remark quite in general in this connection
that the different forms in which measure is realized
also belong to different spheres of natural reality.
The complete, abstract indifference of developed measure,
that is, of its laws, can only be found in the sphere of mechanism
where concrete corporeity is only abstract matter itself;
the qualitative differences of this matter are
of an essentially quantitative nature;
space and time are nothing but pure externalities,
and the aggregates of matters, the masses, the intensity of weight,
are determinations which are just as external
and have their proper determinateness in the quantitative element.
On the other hand, in physical things but even more so in the organic,
this quantitative determinateness of abstract materiality is
already disturbed by the multiplicity
and consequently the conflict of qualities.
And the thus ensuing conflict is not just one of qualities as such,
but measure itself is subordinated here to higher relations
and its immanent development is reduced rather to
the simple form of immediate measure.
The limbs of the animal organism have a measure
which, as a simple quantum, stands
in a ratio to the other quanta of the other limbs;
the proportions of the human body are the fixed ratios of such quanta,
and the science of nature still has far to go in
discovering anything about the link that connects
these magnitudes with the organic functions
on which they are entirely dependent.
But the closest example of the reduction of an immanent
measure to a merely externally determined magnitude is motion.
In the heavenly bodies, motion is free motion,
one which is only determined by the concept from which alone,
consequently, its magnitudes equally depend (see above);
but in the organic body this free motion is reduced
to one which is arbitrary or mechanically regular,
that is, to one which is totally abstract and formal.

And in the realm of spirit there is even less of
a characteristic, free development of measure to be found.
For instance, it is obvious that a
republican constitution like the Athenian,
or an aristocratic constitution mixed with democracy,
is possible only in a state of a certain size;
it is also obvious that in civil society
the multitudes of individuals who belong to
the different occupations stand in a certain ratio to each other.
But none of this yields either laws of measures or typical forms of it.
In the spiritual realm as such there are indeed
distinctions of intensity of character,
strength of imagination, sensations, representations, and so on;
but in determining them one cannot go past this indefinite duo of
“strength” and “weakness.”
To see how lame and totally empty ultimately turn out to be
the so-called laws which have been established for
the relation of strength and weakness in sensations,
representations, and so on, one need only look at
the psychologies that busy themselves with just such matters.
